### 2017-03-20 paper/modes.tex
pass on modes

@@

This paper focuses at lower-level parallelism, which can be enabled
through build systems and testing frameworks.  Given the proliferation
of multi-core machines it is not surprising that testing frameworks
provide today support for parallel test execution (e.g.,
JUnit~\cite{junit-org}, TestNG~\cite{testng}, and NUnit~\cite{nunit}).
In the following we elaborate relevant features of build systems and
testing frameworks.  We focused on Java, Maven, and JUnit but the
discussion can be generalized to other language and tools.

@@

Several build systems support the execution of test suites in forked
processes running a Java virtual machine (JVM).  Each spawned process
runs an individual test suite and the builder outputs the aggregated
reports. When the testing framework finishes the execution, the forked
process can be reused for the next test suite in the queue. In this
level, the number of forked processes must be proportional to the
number of available CPUs otherwise, the high degree of concurrency
may affect the execution performance.  For instance, if the underlying
machine has four CPUs available, the builder could be configured to
spawn one process on each core to favor parallelism (in contrast to
several processes competing for a single core).

@@

\newcommand{\Seq}{L0}
\newcommand{\ParClassSeqMeth}{L1}
\newcommand{\SeqClassParMeth}{L2}
\newcommand{\ParClassParMeth}{L3}

\begin{itemize}
\item
    \textbf{Sequential (\Seq).}~This configuration corresponds to the
    sequential execution of test classes and, for each class,
    sequential execution of their methods.  The execution order (of
    classes and methods) is determined by the testing framework (\eg,
    the order of methods returned by the Java Reflection
    API)~\cite{junit-test-order}.
\item
    \textbf{Sequential classes; parallel methods
      (\ParClassSeqMeth{}).}~This configuration corresponds to the
    sequential execution of the test classes assigned to the JVM.
    However, within each class, test methods run in parallel.
\item
    \textbf{Parallel classes; sequential methods
      (\SeqClassParMeth{}).}~This configuration corresponds to the
    parallel execution of test classes with sequential execution of
    their test methods.
\item
    \textbf{Parallel methods (\ParClassParMeth).}~This configuration
    corresponds to the parallel execution of any test method from any
    class.
\end{itemize}

\Seq{} is the default configuration for JUnit.  It does not use
parallelism, in constrast to the other execution modes.  Considering
use of parallelism, \ParClassParMeth{} is the most agressive mode.
\ParClassSeqMeth{} and \SeqClassParMeth{} run sequentially either
classes or methods.  Intuititvely, fixing one of these dimensions
enables developers to better organize their test suites as to avoid
dataraces~\Fix{cite}.  JUnit implements these modes through a set of
customized classes (e.g., \CodeIn{ParallelComputer}).

\subsubsection*{Example}~ \Fix{The list below shortly describes
the choices to control parallelism within a JVM.}

\begin{figure}[h!]
\centering
\scriptsize
\lstset{
    escapeinside={@}{@},
    numbers=left,xleftmargin=1em,frame=single,framexleftmargin=0.5em,
    basicstyle=\ttfamily\scriptsize, boxpos=c, numberstyle=\tiny,
    morekeywords={parallel, threadCount, perCoreThreadCount},
    deletekeywords={true}
}
\begin{lstlisting}
<plugin>
    <groupId>org.apache.maven.plugins</groupId>
    <artifactId>maven-surefire-plugin</artifactId>
    <configuration>
        <parallel>methods</parallel>
        <threadCount>2</threadCount>
        <perCoreThreadCount>true</perCoreThreadCount>
    </configuration>
</plugin>
\end{lstlisting}
\caption{\label{fig:surefire} Example of Maven configuration: test
  methods run in parallel (\CodeIn{parallel} parameter) using two
  threads (\CodeIn{threadCount} parameter) per available core (flag
  \CodeIn{perCoreThreadCount}).}
\end{figure}

>>>>

\newcommand{\Seq}{L0}
\newcommand{\SeqClassParMeth}{L1}
\newcommand{\ParClassSeqMeth}{L2}
\newcommand{\ParClassParMeth}{L3}
\newcommand{\Fork}{F}
\newcommand{\ForkSeq}{\Fork{}\Seq{}}
\newcommand{\ForkParMeth}{\Fork{}\SeqClassParMeth{}}

@@

This paper focuses at lower-level parallelism, which can be enabled
through build systems and testing frameworks.  Given the proliferation
of multi-core machines it is not surprising that testing frameworks
provide today support for parallel test execution (e.g.,
JUnit~\cite{junit-org}, TestNG~\cite{testng}, and NUnit~\cite{nunit}).
In the following we elaborate relevant features of and testing
frameworks build systems.  We focused on Java, Maven, and JUnit but
the discussion can be generalized to other language and tools.

@@

@@

\begin{itemize}
\item
    \textbf{Sequential (\Seq).}~This configuration corresponds to the
        default behaviour in a test execution. No parallelism is
        involved.
\item
    \textbf{Sequential classes; parallel methods
        (\SeqClassParMeth{}).}~At this level of parallelism, the
        testing framework uses threads to run methods in parallel from
        the assigned test class. Each test class runs sequentially.
\item
    \textbf{Parallel classes; sequential methods
        (\ParClassSeqMeth{}).}~~At this level of parallelism, the
        testing framework uses threads to run classes in parallel and
        each test method runs sequentially within each thread.
\item
    \textbf{Parallel classes; Parallel methods
        (\ParClassParMeth).}~This configuration corresponds
        conceptually to the union of \ParClassSeqMeth{} and
        \SeqClassParMeth{}: the testing framework runs test methods
        from several classes in parallel.
\end{itemize}

Testing frameworks usually execute tests sequentially (\Seq{}) and may
allow the developer to use an ordering criteria (\eg, test name).  For
the \SeqClassParMeth{} setting, \Fix{elaborate and give example...}.
For the \ParClassSeqMeth{} setting, \Fix{elaborate and give
example...}. Finally, the \ParClassParMeth{} setting
\Fix{...elaborate...}.

Several build systems support the execution of test classes in forked
processes running a Java virtual machine (JVM). Each spawned process
runs a test class at time. In addition, the builder can reuse these
processes to execute the next test classes in the queue according to
the developer's preference. For instance, one might want the builder
to spawn a new process for each test class to ensure maximum
isolation.  To exploit parallelism at this level, the number of forked
processes must be proportional to the number of available CPUs,
otherwise, the high frequency of context-switching and concurrency
(\ie, processes competing for a CPU time-slice) may affect the
performance.  For instance, if the underlying machine has four CPUs
available, the builder could be configured to spawn one process on
each core in contrast to several processes competing for a single
core.

Builder-level parallelism can also complement some testing framework
parallelism levels. For instance, \Fix{elaborate...}.  However, some
settings may not take effect since only one test class can run in a
forked process at time. For instance, \Fix{elaborate...}. To the best
of our knowledge, we are unaware of any build system capable of
running multiple classes at time within a forked process.

In this paper, we use the prefix ``\Fork{}'' to indicate that a given
testing configuration has the feature of spawning JVMs to run test
classes enabled. The list below shortly describes the choices to
control parallelism:

\begin{itemize}
\item
    \textbf{Forked classes; sequential methods (\ForkSeq).}~\Fix{todo}
\item
    \textbf{Forked classes; parallel methods (\ForkParMeth).}~\Fix{todo}
\end{itemize}

\subsubsection*{Illustration}~\Fix{elaborate}

\begin{figure}[h!]
\centering
\scriptsize
\lstset{
    escapeinside={@}{@},
    numbers=left,xleftmargin=1em,frame=single,framexleftmargin=0.5em,
    basicstyle=\ttfamily\scriptsize, boxpos=c, numberstyle=\tiny,
    morekeywords={parallel, threadCount, perCoreThreadCount,
    forkCount, reuseFork},
    deletekeywords={true}
}
\begin{lstlisting}
<plugin>
    <groupId>org.apache.maven.plugins</groupId>
    <artifactId>maven-surefire-plugin</artifactId>
    <configuration>
        <forkCount>1C</forkCount>
        <reuseFork>true</true>
        <parallel>methods</parallel>
        <threadCount>1</threadCount>
        <perCoreThreadCount>true</perCoreThreadCount>
    </configuration>
</plugin>
\end{lstlisting}
    \caption{\label{fig:surefire} Maven configured with an
    \Fork{}\SeqClassParMeth{} setting: test methods run in parallel
    (\CodeIn{parallel} parameter) using two threads
    (\CodeIn{threadCount} parameter) per available core (flag
    \CodeIn{perCoreThreadCount}).}
\end{figure}
